---
title: "Final Homework Code 1"
author: "Paige Becker"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged 
    theme: cerulean 
    toc: yes
    toc_float: 
      collapsed: true
      
    

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Reflection  
There are times where I got incredibly frustrated. Coding doesn't come naturally to me, and I don't get that Ah Ha! moment when I do something right. I know a lot of coders feel that, but in my mind, I just get even more frustrated that it took me so long to get something right. However, I did enjoy sticking our code into an RMarkdown. I *love* the knitted files. In my initial code, I couldn't figure out how to use 'str_split()', and Sofia explained it very well in her comments. However, in the end, I decided to keep my code the same because I know coming back to it in a year, I will still understand what I did. I did really appreciate that she took the time to do that. Another thing I found difficult when coding was staying focused. I would finally complete one thing after working on it for awhile, then I'd have to take a massive break. I do get very overwhelmed when coding and I hope to alleviate some of that this semester. The kind comments from my peers did help me gain some confidence in my code, and I think peer commentary from here on out will be great motivation to give myself grace. Julianna also pointed out that I was missing a subgroup in my primate taxonomy which was incredibly helpful, and I applaude her for reading my code so closely.  

Sofia did a great job commenting on every line of code explaining what was happening. It made it very easy to follow. I tried to implement that into my final copy of code. Julianna did a great job of including how she was feeling after each challenge. I did a very similar thing in my code.   


# Challenge 1

### Creating a vector of 1 word character strings comprised of the first line of *Ulysses* by James Joyce

```{r}
u <- c("stately", "plump", "Buck" ,"Mulligan", "came", "from" ,"the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
u
```

#### Extracting every third element of my vector

```{r}
extracted_u <- u[seq(from= 3, to =22, by =3)] #Starting at word 3 to word 22 (I looked at the environment to see how many elements their were) and went by every 3. 
extracted_u
```
I struggled using 'str_split()'. I messed around with it for about 2 hours before I decided to just use seq(). I will ask my peers about it on Thursday. (However, after further inspection, I realized 'str_split()' would be used to split the original vector into several elements instead of extracting elements, but I typed all the words in before realizing that...)  

I think it's okay that I typed in all the words instead of using 'str_split'. Like Prof. Schmitt said, some code is more efficient, and some code is easier to understand. There's no "better" code. If I were to look back on this in a year, I would still understand what it is doing and that is good enough for me.  


# Challenge 2  
### Creating an example matrix
```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE) #creating a matrix with numbers 1-80 numbered by column, with 8 rows and 10 columns
m

```

#### Extracting row 2,3,and 6
```{r}
x <- m[c(2,3,6),] #if you add a c() you can extract multiple rows/columns in this format
x
```
[TutorialKart](https://www.tutorialkart.com/r-tutorial/r-get-multiple-rows-of-matrix/#gsc.tab=0) helped me figure out how to extract *multiple* rows from a matrix.  
This is the structure for extracting information from a matrix:  
**matrixname[row,column]**  

#### Extracting rows 6-8
```{r}
x <- m[c(6:8),]
x
```
#### Extracting elements from row 2, column 2 to row 6, column 9
```{r}
x <- m[2:6,2:9] #adding : lets you extract an elements from one number to what ever other position you want in the matrix
x
```
I am finding that using : is getting more and more intuitive. When you want every element in between 2 locations you can use : in several different functions.  

# Challenge 3

### Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1
```{r}
a <- array(data = 400:1, dim = c(5, 5, 4, 4)) #numbers are 400:1 because we want it in descending order dim= are the dimensions
a
```

Given this matrix, what would the following return?  
a[1, 1, 1, 2]  
395  

###### Checking work  
```{r}
a[1, 1, 1, 2]
```
Ok I was very wrong. I realized I mixed up the locations of the rows and columns. I assumed it was labelled [levels, row, column] but it's actually **[row, column, levels]**. (Albeit, the 4-D concept is a little confusing for me, but idk if that's a coding problem or a brain problem)  

a[2, 3, 2, ]  
364, 264, 164, 64  

###### Checking work  
```{r}
a[2, 3, 2, ] 
```
Success! Although, I am looking at the actual matrix to find numbers instead of doing the math for it.   

a[1:5, 1:5, 3, 3]  
[,1] [,2] [,3] [,4] [,5]  
[1,]  150  145  140  135  130  
[2,]  149  144  139  134  129  
[3,]  148  143  138  133  128  
[4,]  147  142  137  132  127  
[5,]  146  141  136  131  126  
This presumably would give us this whole slice.  

###### Checking work  
```{r}
a[1:5, 1:5, 3, 3] 
```

Yay!  

# Challenge 4  
### Create a list of primate taxonomy  
```{r}
#Setting Up Haplorhini
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae") 
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
Tarsioidea <- c("Tarsiida")
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini") #naming the subgroups under Anthro (I am not sure why they don't automatically get named though)
Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea","Tarsioidea") #naming sub groups under Haplo

#Setting up Strepsirhini
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")

#Combining both
Primates <- list(Haplorhini,Strepsirhini)
names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates


```
Okay, I think I did this successfully, but it's not very pretty. I'd love to learn how to make it look prettier. I started creating the smallest group to the biggest. You can fit the smaller groups inside the bigger ones and esentially stack them that way. I am sure you could create the taxonomy starting from biggest and going to smallest, but as a beginner coder, this way was easier to visualize and comprehend.  

# Practicing with Data Frames  
### Not a challenge, but I thought I'd include it
### Load in data
```{r}
df <- read.csv("C:/Users/pbeck/Desktop/Boston University/Fall 2023/BI 588/AN588-Week-2/random_people", sep = ",", header = TRUE, stringsAsFactors = FALSE) #load data in
#Print select columns
df[, c(1, 3, 4, 11, 12)]
```
### Add additional rows
```{r}

df <- cbind(df, id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16,
    17, 18, 19, 20))
df <- cbind(df, school = c("BU", "BU", "Harvard", "Harvard", "BU", "Boston College",
    "MIT", "BU", "BU", "UMASS Boston", "Harvard", "BU", "Boston College", "BU",
    "Harvard", "MIT", "Harvard", "BU", "UMASS Boston", "Harvard"))
df
```
### Filtering to get specific subsets of data  
This is getting every student who's school is BU
```{r}


new_df <- df[df$school == "BU", ]
new_df
```
Or if we only want the names of people who's school is BU:
```{r}
new_df <- df[df$school == "BU", c("name.last", "name.first", "school")]
new_df
```

# Challenge 5  
### Create 5x3 matrix called m3  
```{r}
m3 <- matrix(data = c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = TRUE)
m3
```
### Coerce the matrix to a data frame  
```{r}
df_m3 <- data.frame(m3)
df_m3
```
### Coerce the second column to be *logical-valued*   
```{r}
df_m3$X2 <- as.logical(df_m3$X2)
str(df_m3)
```
This seemed to work. I did [Google](https://technicaljockey.com/r-programming-training/coercion-in-r/) the 'as.logical()' function. Who knew there's a bunch! 'as.integer()', 'as.character()', 'as.factor()', 'as.numeric()' etc.   


### Coerce the third column to be *factor-valued*  
```{r}
df_m3$X3 <- as.factor(df_m3$X3)
str(df_m3)
```

