---
title: "Prettified Module 4 Challenges"
author: "Paige Becker"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Challenge 1

### Creating a vector of 1 word character strings comprised of the first line of *Ulysses* by James Joyce

```{r}
u <- c("stately", "plump", "Buck" ,"Mulligan", "came", "from" ,"the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
u
```

#### Extracting every third element of my vector

```{r}
extracted_u <- u[seq(from= 3, to =22, by =3)]
extracted_u
# I initially tried to use the extracted command but I think it didn't work because I made the sentence a string. It's the same number of commands but I think it might make your life a little easier if you used str_split() like you said in your comment below. I used the command str_split(v2, " ") so that it would split at every space in the sentence and then a similar command to what you used, unlist(v2)[seq(from = 1, to = 22, by = 3)], to then extract every third word. 
```
I struggled using str_split(). I messed around with it for about 2 hours before I decided to just use seq(). I will ask my peers about it on Thursday. (However, after further inspection, I realized str_split() would be used to split the original vector into several elements instead of extracting elements, but I typed all the words in before realizing that...)

# Challenge 2  
### Creating an example matrix
```{r}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m

```

#### Extracting row 2,3,and 6
```{r}
x <- m[c(2,3,6),]
x
```
[TutorialKart](https://www.tutorialkart.com/r-tutorial/r-get-multiple-rows-of-matrix/#gsc.tab=0) helped me figure out how to extract *multiple* rows from a matrix.  
**matrixname[row,column]**  

#### Extracting rows 6-8
```{r}
x <- m[c(6:8),]
x
```
#### Extracting elements from row 2, column 2 to row 6, column 9
```{r}
x <- m[2:6,2:9]
x
```
I am finding that using : is getting more and more intuitive. When you want every element in between 2 locations you can use : in several different functions. 

# Challenge 3

### Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1
```{r}
a <- array(data = 400:1, dim = c(5, 5, 4, 4))
a
```

Given this matrix, what would the following return?  
a[1, 1, 1, 2]  
395  

###### Checking work  
```{r}
a[1, 1, 1, 2]
```
Ok I was very wrong. I realized I mixed up the locations of the rows and columns. I assumed it was labelled [levels, row, column] but it's actually **[row, column, levels]**. (Albeit, the 4-D concept is a little confusing for me, but idk if that's a coding problem or a brain problem)  

a[2, 3, 2, ]  
364, 264, 164, 64  

###### Checking work  
```{r}
a[2, 3, 2, ] 
```
Success! Although, I am looking at the actual matrix to find numbers instead of doing the math for it.   

a[1:5, 1:5, 3, 3]  
[,1] [,2] [,3] [,4] [,5]  
[1,]  150  145  140  135  130  
[2,]  149  144  139  134  129  
[3,]  148  143  138  133  128  
[4,]  147  142  137  132  127  
[5,]  146  141  136  131  126  
This presumably would give us this whole slice.  

###### Checking work  
```{r}
a[1:5, 1:5, 3, 3] 
```

Yay!  

# Challenge 4  
### Create a list of primate taxonomy  
```{r}
#Setting Up Haplorhini
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Haplorhini <- list(Anthropoidea)
names(Haplorhini) <- list("Anthropoidea")

#Setting up Strepsirhini
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")

#Combining both
Primates <- list(Haplorhini,Strepsirhini)
names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates


```
Okay, I think I did this successfully, but it's not very pretty. I'd love to learn how to make it look prettier.  

# Practicing with Data Frames  
### Not a challenge, but I thought I'd include it
### Load in data
```{r}
df <- read.csv("C:/Users/pbeck/Desktop/Boston University/Fall 2023/BI 588/AN588-Week-2/random_people", sep = ",", header = TRUE, stringsAsFactors = FALSE) #load data in
#Print select columns
df[, c(1, 3, 4, 11, 12)]
```
### Add additional rows
```{r}

df <- cbind(df, id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16,
    17, 18, 19, 20))
df <- cbind(df, school = c("BU", "BU", "Harvard", "Harvard", "BU", "Boston College",
    "MIT", "BU", "BU", "UMASS Boston", "Harvard", "BU", "Boston College", "BU",
    "Harvard", "MIT", "Harvard", "BU", "UMASS Boston", "Harvard"))
df
```
### Filtering to get specific subsets of data  
This is getting every student who's school is BU
```{r}


new_df <- df[df$school == "BU", ]
new_df
```
Or if we only want the names of people who's school is BU:
```{r}
new_df <- df[df$school == "BU", c("name.last", "name.first", "school")]
new_df
```

# Challenge 5  
### Create 5x3 matrix called m3  
```{r}
m3 <- matrix(data = c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = TRUE)
m3
```
### Coerce the matrix to a data frame  
```{r}
df_m3 <- data.frame(m3)
df_m3
```
### Coerce the second column to be *logical-valued*   
```{r}
df_m3$X2 <- as.logical(df_m3$X2)
str(df_m3)
```
This seemed to work. I did [Google](https://technicaljockey.com/r-programming-training/coercion-in-r/) the as.logical() function. Who knew there's a bunch! as.integer(), as.character(), as.factor(), as.numeric etc.   


### Coerce the third column to be *factor-valued*  
```{r}
df_m3$X3 <- as.factor(df_m3$X3)
str(df_m3)
```

